---
title: "eds212-day3-activities"
author: "Julia Parish"
date: "8/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

```

## **Task 1**
- Create an object called pm2_5 with a value of 48 (representing Particulate Matter 2.5, an indicator for air quality, in μgm3 

```{r}
pm2_5 <- 48

```
- Write an if - else if - else statement that returns “Low to moderate risk” if pm2_5 (for Particulate Matter 2.5) is less than 100, “Unhealthy for sensitive groups” if PM 2.5 is 100 <= pm2_5 < 150, and “Health risk present” if PM 2.5 is >= 150.
```{r}
if (pm2_5 < 100) {
  print("Low to Moderate risk")
} else if (pm2_5 <= 100 & pm2_5 < 150) {
  print("Unhealthy for sensitive groups")
} else if (pm2_5 >= 150) {
  print("Health risk present")
}

```
- Test by changing the value of your pm2_5 object and re-running your statement to check.

```{r}
pm2_5 <- 155

if (pm2_5 < 100) {
  print("Low to Moderate risk")
} else if (pm2_5 <= 100 & pm2_5 < 150) {
  print("Unhealthy for sensitive groups")
} else if (pm2_5 >= 150) {
  print("Health risk present")
}

```
## **Task 2**
- Store the string “blue whale” as an object called species. Write an if statement that returns “You found a whale!” if the string “whale” is detected in species, otherwise return nothing. 

```{r}
species <- c("blue whale")

if (str_detect(species, pattern = "whale")) {
  print("You found a whale!")
}

```
### Test by changing the species string & re-running to see output.
```{r}
species <- c("rabbit")

if (str_detect(species, pattern = "whale")) {
  print("You found a whale!")
}
```
# **Task 3**
Create a vector stored as max_airtemp_c with a value of 24.1. Write an if else statement that will print “Temperature too high” if max_airtemp_c is greater than 27, or “Temperature OK” if temperature is less than or equal to 27.
```{r}
max_airtemp_c <- 24.1

if (max_airtemp_c > 27) {
  print("Temperature too high")
} else if (max_airtemp_c <= 27) {
  print("Temperature OK")
}

```
## **Task 4**
- Store the base price of a burrito as base_burrito with a value of 6.50. Store main_ingredient with a starting string of “veggie.” Write a statement that will return the price of a burrito based on what a user specifies as “main_ingredient” (either “veggie”, “chicken” or “steak”) given the following:
- *A veggie burrito is the cost of a base burrito*
- *A chicken burrito costs 3.00 more than a base burrito*
- *A steak burrito costs 3.25 more than a base burrito*

```{r}
base_burrito <- 6.50
main_ingredient <- c("veggie", "chicken", "steak")

burrito_prices <- vector(mode = "numeric", length = length(main_ingredient))

for (i in seq_along(main_ingredient)) {
  if (main_ingredient[i] == "veggie") {
    burrito_price <- base_burrito
  } else if (main_ingredient[i] == "chicken") {
    burrito_price <- base_burrito + 3
  } else if (main_ingredient[i] == "steak") {
    burrito_price <- base_burrito + 3.25
  }
  burrito_prices[[i]] <- burrito_price
}

burrito_prices
```
# **Task 5**
- Create a new vector called fish that contains the values 8, 10, 12, 23 representing counts of different fish types in a fish tank (goldfish, tetras, guppies, and mollies, respectively). Write a for loop that iterates through fish, and returns what proportion of total fish in the tank are that species. Assume that these counts represent all fish in the tank.

```{r}
fish <- c(8, 10, 12, 23)
species <- c("goldfish", "tetras", "guppies", "mollies")

fish_prop <- vector(mode = "numeric", length = length(fish))

for (i in seq_along(fish)) {
  a <- fish[i] / sum(fish) * 100
  fish_prop[i] <- a
} 

fish_prop

names(fish_prop) <- species

fish_prop

```

## **Task 6**
- There is an existing vector in R called month.name that contains all month names (just try running month.name in the Console to check it out). Write a for loop that iterates over all months in month.name and prints “January is month 1,” “February is month 2”, etc.
Hint: you can index values in the month.name vector just like you would any other vector (e.g., try running month.name[5]).

```{r}

month.name[10]

for (i in seq_along(month.name)) {
  print(paste(month.name[i], "is month", i))
}

```
